### 2022-03-15 paper/localpsf.tex
restructuring paper

\section{Preliminaries}

@@

%The ellipsoid $E_x$ is not immediately available. 
%(recall $\gdim$ is the spatial dimension of $\Omega$).
The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}. Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$. But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$, and we will need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ at all points $x$. Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions. Specifically, let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$, and let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
\begin{equation*}
C(x) := 1, \qquad
L^i(x) := x^i, \qquad
Q^{ij}(x) = x^i x^j.
\end{equation*}
The following results (derived in Appendix \ref{app:proofs}) follow straightforwardly from the definitions:
\begin{subequations}
	\label{eq:vol_mean_var}
	\begin{align}
	\spatialvol =& \left(\Aop^T C\right)^* \\
	\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
	\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
	\end{align}
\end{subequations}
for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$. Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.

@@

\subsection{Local mean-displacement invariance}
\label{eq:local_mean_displacement_invariance}

@@

\subsection{Finite element discretization of the integral kernel}
\label{sec:finite_element_kernel}

@@

%We define the finite element approximation of the kernel, $\Phi^h \approx \Phi$, as follows:
The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
\begin{equation}
\label{eq:defn_of_Akerpcmesh}
\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
\end{equation}
%The kernel $\Phi^h$ is the interpolation of $\Phi$ onto $V_h \otimes V_h$. Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$. 
Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
\begin{equation}
\label{eq:kernel_representation_pcmesh}
(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
\end{equation}
%While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the matrix of kernel entries evaluated at pairs of Lagrange nodes:
\begin{equation}
\label{eq:Akerpcmat_entries}
\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
\end{equation}
and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the finite element mass matrix which has entries
\begin{equation*}
\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
\end{equation*}
%Matrix entries $\Phi(p_i, p_j)$ are expensive to compute because $\Aop$ is only available matrix-free.
%Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
%\begin{equation*}
%	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
%\end{equation*}
Our matrix representation of $\Aop$ is therefore given by
\begin{equation}
\label{eq:Amatpc_defn}
\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
\end{equation}
Note that $\mathbf{A}$ is not a Galerkin projection of $\Aop$ onto $V_h$ because of interpolation error in the kernel approximation $\Phi^h \approx \Phi$.

%We neglect this interpolation error because it is the same magnitude as the finite element approximation error. 
%Note that the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are dense $N \times N$ matrix that only exist in theory. Building and storing these matrices is computationally intractible. Computing individual matrix entries $\mathbf{A}_{ij}$ and $\mathbf{\Phi}_{ij}$ is computationally expensive.

@@

Once the impulse response batches $\eta_b$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}). The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
%Specifically, we make the assumption that
%\begin{equation}
%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
%\end{equation}
%if $x_i$ is close to $x$. This is shown in Figure \ref{fig:mean_displacement_invariance}. Motivation for this assumption is discussed in Section \ref{eq:local_mean_displacement_invariance}. 
We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$. The weights are determined by radial basis function interpolation.

>>>>

\section{Key concepts}
\label{sec:prerequisites}

In this section we present three key concepts that our method is based on. The method itself will be presented in Section \ref{sec:method}. First, we define moments of the impulse responses $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}). Our method will use impulse response moments to estimate the supports of the functions $\phi_x$. Second, we describe an impulse response approximation that we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}). Our method will interpolate impulse responses using this approximation. Third, we discuss how discretization of continuous functions by finite elements leads to matrix discretizations $\mathbf{A}$ and $\mathbf{\Phi}$ of $\Aop$ and $\Phi$, respectively (Section \ref{sec:finite_element_kernel}). The goal of our method is to compute H-matrix approximations of these matrices. 
%While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead with minor modifications.

@@

%The ellipsoid $E_x$ is not immediately available. 
%(recall $\gdim$ is the spatial dimension of $\Omega$).
The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}. Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$. But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$, and our method will need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ at all points $x$. Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions. Specifically, let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$, and let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
\begin{equation*}
C(x) := 1, \qquad
L^i(x) := x^i, \qquad
Q^{ij}(x) = x^i x^j.
\end{equation*}
The following results (derived in Appendix \ref{app:proofs}) follow straightforwardly from the definitions:
\begin{subequations}
	\label{eq:vol_mean_var}
	\begin{align}
	\spatialvol =& \left(\Aop^T C\right)^* \\
	\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
	\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
	\end{align}
\end{subequations}
for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$. Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.

@@

\subsection{Local mean-displacement invariance}
\label{sec:local_mean_displacement_invariance}

@@

\subsection{Finite element discretization}
\label{sec:finite_element_kernel}

@@

%We define the finite element approximation of the kernel, $\Phi^h \approx \Phi$, as follows:
The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
\begin{equation}
\label{eq:defn_of_Akerpcmesh}
\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
\end{equation}
%The kernel $\Phi^h$ is the interpolation of $\Phi$ onto $V_h \otimes V_h$. Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$. 
Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
\begin{equation}
\label{eq:kernel_representation_pcmesh}
(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
\end{equation}
%While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
\begin{equation}
\label{eq:Akerpcmat_entries}
\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
\end{equation}
and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the finite element mass matrix which has entries
\begin{equation*}
\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
\end{equation*}
%Matrix entries $\Phi(p_i, p_j)$ are expensive to compute because $\Aop$ is only available matrix-free.
%Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
%\begin{equation*}
%	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
%\end{equation*}
Our matrix representation of $\Aop$ is therefore given by
\begin{equation}
\label{eq:Amatpc_defn}
\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
\end{equation}
Note that $\mathbf{A}$ is not the Galerkin projection of $\Aop$ onto $V_h$ because of the approximation $\Phi^h \approx \Phi$.
%because of the approximation $\Phi^h \approx \Phi$. 
Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the operator action $v=\Aop u$.

%The discrete version of the operator action $v = \Aop u$ is a mapping of vectors $\mathbf{u} \rightarrow \mathbf{v}$, which is done through an implicit computational procedure. Typically the implicitly defined matrix that performs this mapping is the Galerkin projection of $\Aop$ onto $V_h$.

%We neglect this error because it is the same magnitude as the error in finite element approximation of functions $u_h \approx u$. 
%Note that the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are dense $N \times N$ matrix that only exist in theory. Building and storing these matrices is computationally intractible. Computing individual matrix entries $\mathbf{A}_{ij}$ and $\mathbf{\Phi}_{ij}$ is computationally expensive.

@@

Once the impulse response batches $\eta_b$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}). The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{sec:local_mean_displacement_invariance}.
%Specifically, we make the assumption that
%\begin{equation}
%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
%\end{equation}
%if $x_i$ is close to $x$. This is shown in Figure \ref{fig:mean_displacement_invariance}. Motivation for this assumption is discussed in Section \ref{eq:local_mean_displacement_invariance}. 
We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$. The weights are determined by radial basis function interpolation.
